Data cleaning and processing script

### 2023-03-20 code/modeling_code/PP_7DMA_model.Rmd
model building

---
title: "VL_7DMA_model"
author: "Leah Lariscy"
date: "2023-03-17"
output: html_document
---

@@

Predictor: Waste water viral load

@@

This is where we will set VL as predictor and 7DMA as outcome

```{r}
recipe_7dma <- recipe(pcr_pos_7dma ~ total_copies, data = train_data)
```

@@

## Use model on the test data

```{r}
aug_test <- augment(fit_7dma, test_data)
aug_test %>% select(pcr_pos_7dma, .pred)
```

@@

```{r}
rmse <- aug_test %>% rmse(truth = pcr_pos_7dma, .pred)
rsq <- aug_test %>% rsq(truth = pcr_pos_7dma, .pred)
VL_7DMA_metrics <- full_join(rmse, rsq)

saveRDS(VL_7DMA_metrics, here("data/modeling_data/metrics/VL_7DMA_performance.RDS"))

@@

```{r}
aug_test %>% ggplot(aes(pcr_pos_7dma, .pred)) +
  geom_point() +
  stat_smooth(method = "lm")
ggsave(here("data/modeling_data/metrics/VL_7DMA_predict.png"))
```

@@

```{r}
aug_residual <- aug_test %>% mutate(residual = pcr_pos_7dma - .pred)
aug_residual %>% ggplot(aes(.pred, residual)) +
  geom_point()

ggsave(here("data/modeling_data/metrics/VL_7DMA_residual.png"))
```

>>>>

---
title: "PP_7DMA_model"
author: "Leah Lariscy"
date: "2023-03-17"
output: html_document
---

@@

Predictor: Percent positive technical replicates per sample

@@

This is where we will set PP as predictor and 7DMA as outcome

```{r}
recipe_7dma <- recipe(case_7dma ~ pos_rate_tr, data = train_data)
```

@@

## Use model on the training data

```{r}
aug_test <- augment(fit_7dma, train_data)
aug_test %>% select(case_7dma, .pred)
```

@@

```{r}
rmse <- aug_test %>% rmse(truth = case_7dma, .pred)
rsq <- aug_test %>% rsq(truth = case_7dma, .pred)
PP_7DMA_metrics <- full_join(rmse, rsq)

saveRDS(PP_7DMA_metrics, here("data/modeling_data/metrics/PP_7DMA_performance.RDS"))

PP_7DMA_metrics

@@

```{r}
aug_test %>% ggplot(aes(case_7dma, .pred)) +
  geom_point() +
  stat_smooth(method = "lm")
ggsave(here("data/modeling_data/metrics/PP_7DMA_predict.png"))
```

@@

```{r}
aug_residual <- aug_test %>% mutate(residual = case_7dma - .pred)
aug_residual %>% ggplot(aes(.pred, residual)) +
  geom_point()

ggsave(here("data/modeling_data/metrics/PP_7DMA_residual.png"))
```

### 2023-03-20 code/modeling_code/data_splitting.Rmd
model building

```{r}
covid_ww_data <- readRDS(here("data/processed_data/final_combined_data.RDS"))
covid_ww_data <- covid_ww_data %>% drop_na()
tibble(covid_ww_data)
```

>>>>

```{r}
covid_ww_data <- readRDS(here("data/processed_data/final_sum_combined_data.RDS"))
#covid_ww_data <- covid_ww_data %>% drop_na()
tibble(covid_ww_data)
```

### 2023-03-20 code/modeling_code/log10VL_7DMA_model.Rmd
model building

---
title: "VL_7DMA_model"
author: "Leah Lariscy"
date: "2023-03-17"
output: html_document
---

@@

Predictor: Waste water viral load

@@

This is where we will set VL as predictor and 7DMA as outcome

```{r}
recipe_7dma <- recipe(pcr_pos_7dma ~ total_copies, data = train_data)
```

@@

## Use model on the test data

```{r}
aug_test <- augment(fit_7dma, test_data)
aug_test %>% select(pcr_pos_7dma, .pred)
```

@@

```{r}
rmse <- aug_test %>% rmse(truth = pcr_pos_7dma, .pred)
rsq <- aug_test %>% rsq(truth = pcr_pos_7dma, .pred)
VL_7DMA_metrics <- full_join(rmse, rsq)

saveRDS(VL_7DMA_metrics, here("data/modeling_data/metrics/VL_7DMA_performance.RDS"))

@@

```{r}
aug_test %>% ggplot(aes(pcr_pos_7dma, .pred)) +
  geom_point() +
  stat_smooth(method = "lm")
ggsave(here("data/modeling_data/metrics/VL_7DMA_predict.png"))
```

@@

ggsave(here("data/modeling_data/metrics/VL_7DMA_residual.png"))
```

>>>>

---
title: "log10VL_7DMA_model"
author: "Leah Lariscy"
date: "2023-03-17"
output: html_document
---

@@

Predictor: log10 wastewater viral load

@@

This is where we will set log10VL as predictor and 7DMA as outcome

```{r}
recipe_7dma <- recipe(pcr_pos_7dma ~ log10_VL, data = train_data)
```

@@

## Use model on the training data

```{r}
aug_test <- augment(fit_7dma, train_data)
aug_test %>% select(pcr_pos_7dma, .pred)
```

@@

```{r}
rmse <- aug_test %>% rmse(truth = pcr_pos_7dma, .pred)
rsq <- aug_test %>% rsq(truth = pcr_pos_7dma, .pred)
log10VL_7DMA_metrics <- full_join(rmse, rsq)

saveRDS(log10VL_7DMA_metrics, here("data/modeling_data/metrics/log10VL_7DMA_performance.RDS"))

log10VL_7DMA_metrics

@@

```{r}
aug_test %>% ggplot(aes(pcr_pos_7dma, .pred)) +
  geom_point() +
  stat_smooth(method = "lm")
ggsave(here("data/modeling_data/metrics/log10VL_7DMA_predict.png"))
```

@@

ggsave(here("data/modeling_data/metrics/log10VL_7DMA_residual.png"))
```

### 2023-03-20 code/modeling_code/log10VL_PP_7DMA_model.Rmd
model building

---
title: "log10VL_7DMA_model"
author: "Leah Lariscy"
date: "2023-03-17"
output: html_document
---

@@

Predictor: log10 wastewater viral load

@@

This is where we will set log10VL as predictor and 7DMA as outcome

```{r}
recipe_7dma <- recipe(pcr_pos_7dma ~ log10_VL, data = train_data)
```

@@

```{r}
rmse <- aug_test %>% rmse(truth = pcr_pos_7dma, .pred)
rsq <- aug_test %>% rsq(truth = pcr_pos_7dma, .pred)
log10VL_7DMA_metrics <- full_join(rmse, rsq)

saveRDS(log10VL_7DMA_metrics, here("data/modeling_data/metrics/log10VL_7DMA_performance.RDS"))

log10VL_7DMA_metrics

@@

```{r}
aug_test %>% ggplot(aes(pcr_pos_7dma, .pred)) +
  geom_point() +
  stat_smooth(method = "lm")
ggsave(here("data/modeling_data/metrics/log10VL_7DMA_predict.png"))
```

@@

ggsave(here("data/modeling_data/metrics/log10VL_7DMA_residual.png"))
```

>>>>

---
title: "log10VL_PP_7DMA_model"
author: "Leah Lariscy"
date: "2023-03-17"
output: html_document
---

@@

Predictor: Waste water viral load

@@

This is where we will set VL as predictor and 7DMA as outcome

```{r}
recipe_7dma <- recipe(pcr_pos_7dma ~ log10_VL + pos_rate_tr, data = train_data)
```

@@

```{r}
rmse <- aug_test %>% rmse(truth = pcr_pos_7dma, .pred)
rsq <- aug_test %>% rsq(truth = pcr_pos_7dma, .pred)
log10VL_PP_7DMA_metrics <- full_join(rmse, rsq)

saveRDS(log10VL_PP_7DMA_metrics, here("data/modeling_data/metrics/log10VL_PP_7DMA_performance.RDS"))

log10VL_PP_7DMA_metrics

@@

```{r}
aug_test %>% ggplot(aes(pcr_pos_7dma, .pred)) +
  geom_point() +
  stat_smooth(method = "lm")
ggsave(here("data/modeling_data/metrics/log10VL_PP_7DMA_predict.png"))
```

@@

ggsave(here("data/modeling_data/metrics/log10VL_PP_7DMA_residual.png"))
```

### 2023-03-20 code/processing_code/processing.qmd
model building

```{r}
final_data <- full_join(all_data, wbe_samp, by = c("date", "facility"))
saveRDS(final_data, here("data/processed_data/final_combined_data.RDS"))
```

>>>>

```{r}
final_data <- full_join(all_data, wbe_samp, by = c("date", "facility")) %>% 
  mutate(log10_VL = log10(total_copies), percent_pos_test = prop_pos*100) %>% 
  drop_na(facility)

saveRDS(final_data, here("data/processed_data/final_combined_data.RDS"))

final_data
```

### Summarize data across treatment facilities then average across targets

Group by: Date, Collection number

Keep: total copies, pos_rate_tr, log10 total copies, % positive tests

```{r}
final_data_sum <- final_data %>% group_by(date, collection_num, target) %>% 
  summarize(sum_total_copies = sum(total_copies),
            pos_rate_tr = mean(pos_rate_tr),
            percent_pos_test = mean(percent_pos_test),
            case_7dma = mean(pcr_pos_7dma)) %>% 
  ungroup(target) %>% 
  summarize(avg_VL = mean(sum_total_copies),
            pos_rate_tr = mean(pos_rate_tr),
            percent_pos_test = mean(percent_pos_test),
            case_7dma = mean(case_7dma)) %>% 
  ungroup() %>% 
  mutate(log10_VL = log10(avg_VL))

saveRDS(final_data_sum, here("data/processed_data/final_sum_combined_data.RDS"))
```
